The Error Estimate for Computing Electrostatic Interaction in the Inhomogeneous and Correlated Molecular Systems: for the Ewald Summation, SPME and Staggered Mesh Ewald.

### 2012-04-25 longRange.hetero/publish/main.tex
using the concept of testing particle.

In this paper, we consider the system with periodic boundary condition.
Suppose it is
composed by $N$ charged particles located at $\v r_1, \v r_2, \cdots,
\v r_N$ with charges $q_1, q_2, \cdots, q_N$, respectively.  We study
the force error of an arbitrary charge $i$ by calculating the
difference between the exact force and calculated force exerting on
the charge $i$, which is called the \emph{error force}
\cite{wang2012} and denoted by $\Delta \v F(\v r_i)$.
% and is denoted by
% $\Delta \v F(\v r_i) = \v F(\v r_i) - \tilde{\v F}(\v r_i)$.
The magnitude of the force computation error 
% error of the force computation
is defined by the \emph{root mean square (RMS)} \emph{error}, which is
the square root of the second moment of the error force: $\mathcal
E(\v r_i) = \sqrt{\langle\vert\Delta\v F(\v r_i)\vert^2\rangle}$.  The
notation $\langle\cdot\rangle$ means ensemble average.  We also
estimate the the ensemble average of the error force (also called
\emph{mean error force}), namely $\langle\Delta\v F(\v r_i)\rangle$,
which plays an important role in the error analysis of the
inhomogeneous systems~\cite{wang2012}.

@@

  If the error force of an arbitrary particle $i$ is

@@

    \Delta \v F(\v r_i) =

@@

    \sum_{j=1}^N\,q_iq_j \v K(\v r_i, \v r_j),

@@

    \langle\Delta\v F(\v r_i)\rangle
    =q_i\, \int_{\mathbb R^3}\v K (\v r_i,\v r')\,\rho_q(\v r')\,\d d\v r',

@@

    \mathcal E^2 (\v r_i) 

@@

    q_i^2\,\int_{\mathbb R^3}\vert\v K(\v r_i,\v r')\vert^2\,\rho_{q^2} (\v r')\,\d d\v r' + 
    q_i^2\,

@@

    \int_{\mathbb R^3}\v K(\v r_i,\v r')\,\rho_{q}(\v r')\,\d d\v r'

@@

    q_i^2\,\int_{\mathbb R^3\times\mathbb R^3}
    \v K(\v r_i,\v r')\cdot
    \v K(\v r_i,\v r'')\,

@@

    \langle\Delta\v F(\v r_i)\rangle

@@

    q_i\bigg\langle \sum_{j=1}^Nq_j\v K(\v r_i,\v r_j)\bigg\rangle\\

@@

    q_i\int_{\mathbb R^3}\v K(\v r_i, \v r')

@@

    q_i\,\int\v K(\v r_i,\v r')\,\rho_q(\v r')\,\d d\v r'.

@@

    \langle\vert\Delta\v F(\v r_i)\vert^2\rangle

@@

    q_i^2\bigg\langle\sum_{j,k}
    q_jq_k\v K(\v r_i, \v r_j)\cdot\v K(\v r_i, \v r_k)\bigg\rangle \\ \nonumber

@@

    q_i^2\bigg\langle\sum_{j=1}^N
    q_j^2\vert\v K(\v r_i, \v r_j)\vert^2

@@

    q_i^2\bigg\langle\sum_{j\neq k}
    q_jq_k \v K(\v r_i, \v r_j)\cdot\v K(\v r_i, \v r_k)

@@

    q_i^2\int_{\mathbb R^3}
    \vert\v K(\v r_i, \v r')\vert^2\rho_{q^2}(\v r')\,\d d\v r'

@@

    q_i^2\int_{\mathbb R^3\times\mathbb R^3}
    \v K(\v r_i , \v r')\cdot\v K(\v r_i , \v r'')\,

@@

    q_i^2\int_{\mathbb R^3}
    \vert\v K(\v r_i , \v r')\vert^2\rho_{q^2}(\v r')\,\d d\v r'

@@

    q_i^2

@@

    \v K(\v r_i , \v r')\rho_{q}(\v r')\,\d d\v r'

@@

    % q_i^2\,[(\v K)^2\ast\rho_{q^2}] (\v r_i) + 
    % q_i^2\,[\v K\ast\rho_{q}]^2 (\v r_i) \\

@@

    % q_i^2\,\int_{\mathbb R^3\times\mathbb R^3}
    % \v K(\v r_i-\v r')\cdot
    % \v K(\v r_i-\v r'')\,

@@

    \v K(\v r_i,\v r_j) = \v K(\v r_i - \v r_j),    

@@

    \langle\Delta\v F(\v r_i)\rangle
    =q_i\, [\v K\ast\rho_q](\v r_i),

@@

    \mathcal E (\v r_i) 

@@

    q_i^2\,[(\v K)^2\ast\rho_{q^2}] (\v r_i) + 
    q_i^2\,[\v K\ast\rho_{q}]^2 (\v r_i) \\\label{eqn:cor-error}

@@

    q_i^2\,\int_{\mathbb R^3\times\mathbb R^3}
    \v K(\v r_i-\v r')\cdot
    \v K(\v r_i-\v r'')\,

@@

  $C_{q^2}(\v r, \v r')$ defined in

@@

  different charges. When they are independent, $C_{q^2}(\v r, \v r')$
  vanishes, otherwise $C_{q^2}(\v r, \v r') \neq 0$.

@@

  The function $\v K(\v r_i, \v r_j)$ is called \emph{the error force

@@

    \mathcal E^2_{\textrm{homo}}(\v r_i) +
    \mathcal E^2_{\textrm{inhomo}}(\v r_i) +
    \mathcal E_{\textrm{correlation}}(\v r_i).

@@

  Without the property $\v K(\v r_i,\v r_j) = \v K(\v r_i - \v r_j)$,
  theorem~\ref{thm:tmp1} calculate the homogeneity and inhomogeneity
  error at a cost of $\mathcal O(N^2\log N)$, and correlation error
  at a cost of $\mathcal O(N^3\log N)$.

@@

very flexible, basically it means the nearby neighbors that are

@@

  C_{q^2}(\v r', \v r'') \approx

@@

  \mathcal E_{\textrm{correlation}}(\v r_i)
  \approx q_i^2\sum_j\sum_{k\in\Omega_j}q_jq_k

@@

  \v K(\v r_i , \v r_j)\cdot\v K(\v r_i , \v r_k)

@@

``nearest neighbors''. In the system of liquid and gas, the
correlation is only short-ranged, so it is possible to systematically
improve the accuracy by considering more and more neighbors as long as

@@

is very strong.  Therefore, in this case, we define the nearest neighbor by the

@@

  \mathcal E_{\textrm{correlation}}(\v r_i)

@@

  q_i^2\sum_{j\in\Omega_{\textrm{O}}}\sum_{k\in\Omega_j}q_jq_k

@@

  q_i^2\sum_{j\in\Omega_{\textrm{H}}}\sum_{k\in\Omega_j}q_jq_k

@@

  \v K(\v r_i , \v r')\cdot

@@

  q_{\textrm{H}}\v K(\v r_i , \v r' + \v s_{\textrm{O}})

@@

  q_{\textrm{O}}\v K(\v r_i , \v r' - \v s_{\textrm{O}})+
  q_{\textrm{H}}\v K(\v r_i , \v r' + \v s_{\textrm{H}})

@@

has the form $\v K(\v r_i , \v r_j) = \v K(\v r_i - \v r_j)$, 

@@

  q_{\textrm{H}}\v K(\v r_i - \v r' + \v s_{\textrm{O}})

@@

  e^{2\pi i\v m\cdot(\v r_i - \v r' + \v s_{\textrm{O}})} 

@@

  q_{\textrm{H}}\,e^{2\pi i\v m\cdot\v s_{\textrm{O}}}

@@

  e^{2\pi i\v m\cdot(\v r_i - \v r')},

@@

  q_{\textrm{O}}\,e^{-2\pi i\v m\cdot\v s_{\textrm{O}}} +
  q_{\textrm{H}}\,e^{2\pi i\v m\cdot\v s_{\textrm{H}}}

@@

  \v K(\v r_i - \v r')\cdot
  (\hat T_{\textrm{O}}\hat{\v K})^{\vee}(\v r_i - \v r')

@@

  \v K(\v r_i - \v r')\cdot
  (\hat T_{\textrm{H}}\hat{\v K})^{\vee}(\v r_i - \v r')

@@

  \Delta\v F_{\textrm{rec}}^{\textrm{ik}}(\v r_i)

@@

  q_i

@@

  A(\v m, \v r_i)\,
  e^{2\pi i\v m\cdot \v r_i}
  \sum_{j\neq i}q_j\,e^{-2\pi i\v m\cdot\v r_j} \\\nonumber

@@

  e^{2\pi i\v m\cdot \v r_i}
  \sum_{j\neq i}q_j\,A(\v m, -\v r_j)\,e^{-2\pi i\v m\cdot\v r_j}\\\label{eqn:ik.error.force.1}

@@

  A(\v m, \v r_i) +
  A(\v m, -\v r_i)

@@

  A(\v m, \v r_i)\,
  e^{2\pi i\v m\cdot \v r_i}
  \sum_{j\neq i}q_j\,e^{-2\pi i\v m\cdot\v r_j} \\\label{eqn:ik.error.force}

@@

  e^{2\pi i\v m\cdot \v r_i}
  \sum_{j\neq i}q_j\,A(\v m, -\v r_j)\,e^{-2\pi i\v m\cdot\v r_j}

@@

  - 2 q_i

@@

  \big] (\v r_i),

@@

  \vert\Delta\v F_{\textrm{rec}}^{\textrm{ik}}(\v r_i)\vert^2

@@

  2q_i^2

@@

  \,\big] (\v r_i)\\\nonumber

@@

  4q_i^2\,

@@

  \big] (\v r_i) \\\label{eqn:ik.error}

@@

  \mathcal E^{\textrm{ik}}_{\textrm{correlation}}(\v r_i)

@@

  \,\big] (\v r_i) \\\nonumber

@@

  \big] (\v r_i) \\\nonumber

@@

  \,\big] (\v r_i) \\

@@

  \big] (\v r_i).

@@

The error force of the analytical differentiation is 
\begin{align}\nonumber
  \Delta\v F_{\textrm{rec}}^{\textrm{ana}}(\v r_i)
  % =\;&
  % q_i
  % \sum_{
  %   \begin{subarray}{c}
  %     \vert m_{\alpha}\vert < K_\alpha/2\\
  %     \v m\neq 0
  %   \end{subarray}}
  % -2f(\v m)\,
  % \nabla_{\v r_i}
  % e^{2\pi i\v m\cdot \v r_i}
  % \sum_{j\neq i}q_j\,
  % A(\v m, -\v r_j)
  % e^{-2\pi i\v m\cdot\v r_j} \\\nonumber
  % & +\,
  % q_i
  % \sum_{
  %   \begin{subarray}{c}
  %     \vert m_{\alpha}\vert < K_\alpha/2\\
  %     \v m\neq 0
  %   \end{subarray}}
  % -2f(\v m)\,
  % \nabla_{\v r_i}
  % [A(\v m, \v r_i)\,
  % e^{2\pi i\v m\cdot \v r_i} ]
  % \sum_{j\neq i}q_j\,
  % e^{-2\pi i\v m\cdot\v r_j} \\\nonumber
  % & +\,
  % q^2_i
  % \sum_{
  %   \begin{subarray}{c}
  %     \vert m_{\alpha}\vert < K_\alpha/2\\\nonumber
  %     \v m\neq 0
  %   \end{subarray}}
  % -2f(\v m)\,
  % A(-\v m, \v r_i)\,e^{2\pi i\v m\cdot \v r_i} \nabla_{\v r_i}e^{2\pi i\v m\cdot\v r_i} \\\nonumber
  % & +\,
  % q^2_i
  % \sum_{
  %   \begin{subarray}{c}
  %     \vert m_{\alpha}\vert < K_\alpha/2\\
  %     \v m\neq 0
  %   \end{subarray}}
  % -2f(\v m)\,
  % e^{-2\pi i\v m\cdot \v r_i}
  % \nabla_{\v r_i} [\,A(\v m, \v r_i) \,e^{2\pi i\v m\cdot\v r_i}\,]  \\\nonumber
  =&\,
  \Delta\v F_{\textrm{rec}}^{\textrm{ik}}(\v r_i)\\\nonumber
  & +\,
  q_i
  \sum_{
    \begin{subarray}{c}
      \vert m_{\alpha}\vert < K_\alpha/2\\
      \v m\neq 0
    \end{subarray}}
  -2f(\v m)\,
  \v B(\v m, \v r_i)\,e^{2\pi i\v m\cdot \v r_i}
  \sum_{j\neq i}q_j\,
  e^{-2\pi i\v m\cdot\v r_j} \\\label{eqn:ana.error.force}
  & +\,
  q^2_i
  \sum_{
    \begin{subarray}{c}
      \vert m_{\alpha}\vert < K_\alpha/2\\
      \v m\neq 0
    \end{subarray}}
  -2f(\v m)\,\v B(\v m, \v r_i) 
\end{align}
With $\v B(\v m, \v r)$ defined by
\begin{align}\nonumber
  \v B(\v m, \v r)
  &=
  \nabla_{\v r}\,A(\v m, \v r)\\\label{eqn:error-B}
  &=
  \sum_{\alpha=1}^3\sum_{l\neq 0}
  Z_{\alpha,l}(\v m)\,
  2\pi i l K_\alpha\v a_\alpha^\ast\,e^{2\pi i l K_\alpha\v a^\ast_\alpha\cdot \v r} 
\end{align}
The last term in Eqn.~\eqref{eqn:ana.error.force} is due to the
self-interaction~\cite{cerutti2009staggered, ballenegger2011removal,
  neelov2010interlaced}, which does not depends on the coordinates of
particles. Therefore, it can be calculated explicitly, and subtracted
from the analytical force during the simulation. The computational cost of
the self-interaction term is very low comparing with the SPME
force calculation. It is shown that the self-interaction  error
dominate in the low charge density regions~\cite{cerutti2009staggered}.
Therefore, we always remove the self-interaction term from
the analytical differentiation.
% , and call it as the original analytical
% differetiation when comparing it with the staggered mesh analytical
% differetiation.

>>>>

In this paper, we consider the system with periodic boundary condition.
Suppose it is
composed by $N$ charged particles located at $\v r_1, \v r_2, \cdots,
\v r_N$ with charges $q_1, q_2, \cdots, q_N$, respectively.
We study the force of a testing particle located at $\v r$, which
feels interactions from all charges $q_1, q_2, \cdots, q_N$
in the system, and does not
exert force on them.
The \emph{error force} is defined by the difference between
the exact force and the calculated
force on the testing particle
\cite{wang2012}, and is denoted by $\Delta \v F(\v r)$.
The magnitude of the force computation error 
is defined by the \emph{root mean square (RMS)} \emph{error}, which is
the square root of the second moment of the error force: $\mathcal
E(\v r) = \sqrt{\langle\vert\Delta\v F(\v r)\vert^2\rangle}$.  The
notation $\langle\cdot\rangle$ means ensemble average.  We also
estimate the the ensemble average of the error force (also called
\emph{mean error force}), namely $\langle\Delta\v F(\v r)\rangle$,
which plays an important role in the error analysis of the
inhomogeneous systems~\cite{wang2012}.

@@

  If the error force of the testing particle with charge $q$ has the form:

@@

    \Delta \v F(\v r) =

@@

    q\sum_{j=1}^N\,q_j \v K(\v r, \v r_j),

@@

    \langle\Delta\v F(\v r)\rangle
    =q\, \int_{\mathbb R^3}\v K (\v r,\v r')\,\rho_q(\v r')\,\d d\v r',

@@

    \mathcal E^2 (\v r) 

@@

    q^2\,\int_{\mathbb R^3}\vert\v K(\v r,\v r')\vert^2\,\rho_{q^2} (\v r')\,\d d\v r' + 
    q^2\,

@@

    \int_{\mathbb R^3}\v K(\v r,\v r')\,\rho_{q}(\v r')\,\d d\v r'

@@

    q^2\,\int_{\mathbb R^3\times\mathbb R^3}
    \v K(\v r,\v r')\cdot
    \v K(\v r,\v r'')\,

@@

    \langle\Delta\v F(\v r)\rangle

@@

    q\bigg\langle \sum_{j=1}^Nq_j\v K(\v r,\v r_j)\bigg\rangle\\

@@

    q\int_{\mathbb R^3}\v K(\v r, \v r')

@@

    q\,\int_{\mathbb R^3}\v K(\v r,\v r')\,\rho_q(\v r')\,\d d\v r'.

@@

    \langle\vert\Delta\v F(\v r)\vert^2\rangle

@@

    q^2\bigg\langle\sum_{j,k}
    q_jq_k\v K(\v r, \v r_j)\cdot\v K(\v r, \v r_k)\bigg\rangle \\ \nonumber

@@

    q^2\bigg\langle\sum_{j=1}^N
    q_j^2\vert\v K(\v r, \v r_j)\vert^2

@@

    q^2\bigg\langle\sum_{j\neq k}
    q_jq_k \v K(\v r, \v r_j)\cdot\v K(\v r, \v r_k)

@@

    q^2\int_{\mathbb R^3}
    \vert\v K(\v r, \v r')\vert^2\rho_{q^2}(\v r')\,\d d\v r'

@@

    q^2\int_{\mathbb R^3\times\mathbb R^3}
    \v K(\v r , \v r')\cdot\v K(\v r , \v r'')\,

@@

    q^2\int_{\mathbb R^3}
    \vert\v K(\v r , \v r')\vert^2\rho_{q^2}(\v r')\,\d d\v r'

@@

    q^2

@@

    \v K(\v r , \v r')\rho_{q}(\v r')\,\d d\v r'

@@

    % q^2\,[(\v K)^2\ast\rho_{q^2}] (\v r) + 
    % q^2\,[\v K\ast\rho_{q}]^2 (\v r) \\

@@

    % q^2\,\int_{\mathbb R^3\times\mathbb R^3}
    % \v K(\v r-\v r')\cdot
    % \v K(\v r-\v r'')\,

@@

    \v K(\v r,\v r') = \v K(\v r - \v r'),    

@@

    \langle\Delta\v F(\v r)\rangle
    =q\, [\v K\ast\rho_q](\v r),

@@

    \mathcal E (\v r) 

@@

    q^2\,[(\v K)^2\ast\rho_{q^2}] (\v r) + 
    q^2\,[\v K\ast\rho_{q}]^2 (\v r) \\\label{eqn:cor-error}

@@

    q^2\,\int_{\mathbb R^3\times\mathbb R^3}
    \v K(\v r-\v r')\cdot
    \v K(\v r-\v r'')\,
    C(\v r', \v r'')\,

@@

  $C(\v r, \v r')$ defined in

@@

  different charges. When they are independent, $C(\v r, \v r')$
  vanishes, otherwise $C(\v r, \v r') \neq 0$.

@@

  The function $\v K(\v r, \v r')$ is called \emph{the error force

@@

    \mathcal E^2_{\textrm{homo}}(\v r) +
    \mathcal E^2_{\textrm{inhomo}}(\v r) +
    \mathcal E_{\textrm{correlation}}(\v r).

@@

  Without the property $\v K(\v r,\v r') = \v K(\v r - \v r')$,
  theorem~\ref{thm:tmp1} calculates the homogeneity and inhomogeneity
  error at a cost of $\mathcal O(N^2)$, and correlation error
  at a cost of $\mathcal O(N^3)$.

@@

very flexible. Basically it means the nearby neighbors that are

@@

  C(\v r', \v r'') \approx

@@

  \mathcal E_{\textrm{correlation}}(\v r)
  \approx q^2\sum_j\sum_{k\in\Omega_j}q_jq_k

@@

  \v K(\v r , \v r_j)\cdot\v K(\v r , \v r_k)

@@

``nearest neighbors''. In the systems
of liquid and gas far from critical point,
the correlation is only short-ranged, so charges are nearest neighbors when
their correlations are considered important.
It is also possible to systematically
improve the approximation by considering more and more neighbors as long as

@@

is important.  Therefore, in this case, we define the nearest neighbor by the

@@

  \mathcal E_{\textrm{correlation}}(\v r)

@@

  q^2\sum_{j\in\Omega_{\textrm{O}}}\sum_{k\in\Omega_j}q_jq_k

@@

  q^2\sum_{j\in\Omega_{\textrm{H}}}\sum_{k\in\Omega_j}q_jq_k

@@

  \v K(\v r , \v r')\cdot

@@

  q_{\textrm{H}}\v K(\v r , \v r' + \v s_{\textrm{O}})

@@

  q_{\textrm{O}}\v K(\v r , \v r' - \v s_{\textrm{O}})+
  q_{\textrm{H}}\v K(\v r , \v r' + \v s_{\textrm{H}})

@@

has the form $\v K(\v r , \v r') = \v K(\v r - \v r')$, 

@@

  q_{\textrm{H}}\v K(\v r - \v r' - \v s_{\textrm{O}})

@@

  e^{2\pi i\v m\cdot(\v r - \v r' - \v s_{\textrm{O}})} 

@@

  q_{\textrm{H}}\,e^{-2\pi i\v m\cdot\v s_{\textrm{O}}}

@@

  e^{2\pi i\v m\cdot(\v r - \v r')},

@@

  q_{\textrm{O}}\,e^{2\pi i\v m\cdot\v s_{\textrm{O}}} +
  q_{\textrm{H}}\,e^{-2\pi i\v m\cdot\v s_{\textrm{H}}}

@@

  \v K(\v r - \v r')\cdot
  (\hat T_{\textrm{O}}\hat{\v K})^{\vee}(\v r - \v r')

@@

  \v K(\v r - \v r')\cdot
  (\hat T_{\textrm{H}}\hat{\v K})^{\vee}(\v r - \v r')

@@

  \Delta\v F_{\textrm{rec}}^{\textrm{ik}}(\v r)

@@

  q

@@

  A(\v m, \v r)\,
  e^{2\pi i\v m\cdot \v r}
  \sum_{j}q_j\,e^{-2\pi i\v m\cdot\v r_j} \\\nonumber

@@

  e^{2\pi i\v m\cdot \v r}
  \sum_{j}q_j\,A(\v m, -\v r_j)\,e^{-2\pi i\v m\cdot\v r_j}\\\label{eqn:ik.error.force.1}

@@

  A(\v m, \v r) +
  A(\v m, -\v r)

@@

  A(\v m, \v r)\,
  e^{2\pi i\v m\cdot \v r}
  \sum_{j}q_j\,e^{-2\pi i\v m\cdot\v r_j} \\\label{eqn:ik.error.force}

@@

  e^{2\pi i\v m\cdot \v r}
  \sum_{j}q_j\,A(\v m, -\v r_j)\,e^{-2\pi i\v m\cdot\v r_j}

@@

  - 2 q

@@

  \big] (\v r),

@@

  \vert\Delta\v F_{\textrm{rec}}^{\textrm{ik}}(\v r)\vert^2

@@

  2q^2

@@

  \,\big] (\v r)\\\nonumber

@@

  4q^2\,

@@

  \big] (\v r) \\\label{eqn:ik.error}

@@

  \mathcal E^{\textrm{ik}}_{\textrm{correlation}}(\v r)

@@

  \,\big] (\v r) \\\nonumber

@@

  \big] (\v r) \\\nonumber

@@

  \,\big] (\v r) \\

@@

  \big] (\v r).

@@

The error force of the analytical differentiation is 
\begin{align}\nonumber
  \Delta\v F_{\textrm{rec}}^{\textrm{ana}}(\v r)
  % =\;&
  % q
  % \sum_{
  %   \begin{subarray}{c}
  %     \vert m_{\alpha}\vert < K_\alpha/2\\
  %     \v m\neq 0
  %   \end{subarray}}
  % -2f(\v m)\,
  % \nabla_{\v r}
  % e^{2\pi i\v m\cdot \v r}
  % \sum_{j\neq i}q_j\,
  % A(\v m, -\v r_j)
  % e^{-2\pi i\v m\cdot\v r_j} \\\nonumber
  % & +\,
  % q
  % \sum_{
  %   \begin{subarray}{c}
  %     \vert m_{\alpha}\vert < K_\alpha/2\\
  %     \v m\neq 0
  %   \end{subarray}}
  % -2f(\v m)\,
  % \nabla_{\v r}
  % [A(\v m, \v r)\,
  % e^{2\pi i\v m\cdot \v r} ]
  % \sum_{j\neq i}q_j\,
  % e^{-2\pi i\v m\cdot\v r_j} \\\nonumber
  % & +\,
  % q^2
  % \sum_{
  %   \begin{subarray}{c}
  %     \vert m_{\alpha}\vert < K_\alpha/2\\\nonumber
  %     \v m\neq 0
  %   \end{subarray}}
  % -2f(\v m)\,
  % A(-\v m, \v r)\,e^{2\pi i\v m\cdot \v r} \nabla_{\v r}e^{2\pi i\v m\cdot\v r} \\\nonumber
  % & +\,
  % q^2
  % \sum_{
  %   \begin{subarray}{c}
  %     \vert m_{\alpha}\vert < K_\alpha/2\\
  %     \v m\neq 0
  %   \end{subarray}}
  % -2f(\v m)\,
  % e^{-2\pi i\v m\cdot \v r}
  % \nabla_{\v r} [\,A(\v m, \v r) \,e^{2\pi i\v m\cdot\v r}\,]  \\\nonumber
  =&\,
  \Delta\v F_{\textrm{rec}}^{\textrm{ik}}(\v r)\\\nonumber
  & +\,
  q
  \sum_{
    \begin{subarray}{c}
      \vert m_{\alpha}\vert < K_\alpha/2\\
      \v m\neq 0
    \end{subarray}}
  -2f(\v m)\,
  \v B(\v m, \v r)\,e^{2\pi i\v m\cdot \v r}
  \sum_{j}q_j\,
  e^{-2\pi i\v m\cdot\v r_j} \\\label{eqn:ana.error.force}
  & +\,
  q^2
  \sum_{
    \begin{subarray}{c}
      \vert m_{\alpha}\vert < K_\alpha/2\\
      \v m\neq 0
    \end{subarray}}
  -2f(\v m)\,\v B(\v m, \v r) 
\end{align}
With $\v B(\v m, \v r)$ defined by
\begin{align}\nonumber
  \v B(\v m, \v r)
  &=
  \nabla_{\v r}\,A(\v m, \v r)\\\label{eqn:error-B}
  &=
  \sum_{\alpha=1}^3\sum_{l\neq 0}
  Z_{\alpha,l}(\v m)\,
  2\pi i l K_\alpha\v a_\alpha^\ast\,e^{2\pi i l K_\alpha\v a^\ast_\alpha\cdot \v r} 
\end{align}
The last term in Eqn.~\eqref{eqn:ana.error.force} is due to the
self-interaction~\cite{cerutti2009staggered, ballenegger2011removal,
  neelov2010interlaced}, which does not depends on the coordinates of
particles. Therefore, it can be calculated explicitly, and subtracted
from the analytical force during the simulation. The computational cost of
the self-interaction term is very low comparing with the SPME
force calculation. It is shown that the self-interaction  error
dominate in the low charge density regions~\cite{cerutti2009staggered}.
Therefore, we always remove the self-interaction term from
the analytical differentiation.
% , and call it as the original analytical
% differetiation when comparing it with the staggered mesh analytical
% differetiation.

Neglecting all high frequency terms, we have the mean error force of
analytical differentiation:
\begin{align}\label{eqn:ana.meanf}
  \langle \Delta\v F_{\textrm{rec}}^{\textrm{ana}}(\v r)\rangle
  \approx
  \langle \Delta\v F_{\textrm{rec}}^{\textrm{ik}}(\v r)\rangle.
\end{align}
The RMS force error is 
\begin{align}\nonumber
  \big\langle
  \vert \Delta\v F_{\textrm{rec}}^{\textrm{ana}}(\v r)\vert^2
  \big\rangle
  \approx&\,
  q^2
  \big[
  \big(
  \sum_{\alpha} \sum_{l\neq 0}
  \vert \v G_{\alpha,l}\vert^2
  \big)
  \ast \rho_{q^2}
  \,\big] (\v r) \\\nonumber
  & +\,
  q^2
  \big[
  \big(
  \sum_{\alpha}\sum_{l\neq 0}
  \vert
  % \v G_{\alpha,l} + 2\pi i\,l K_\alpha F_{\alpha,l} \v a_\alpha^\ast\vert^2
  \v G_{\alpha,l} + \v F_{\alpha,l}\vert^2
  \big)
  \ast \rho_{q^2}
  \big]
  (\v r) \\ \nonumber
  & +\,
  4q^2\,
  \big[
  \big(
  \sum_{\alpha} \sum_{l\neq 0}  
  \v G_{\alpha,l}
  \big)^2
  \ast \rho_{q^2}\,
  \big] (\v r) \\\label{eqn:ana.error}
  & +\,
  \big\langle \Delta\v F_{\textrm{rec}}^{\textrm{ana}}(\v r)\big\rangle ^2
  % &+\,
  % q^4 \sum_{\alpha}\sum_{l\neq 0}
  % 4\pi^2l^2K_\alpha^2\vert\v a_{\alpha}^\ast\vert^2[C_{\alpha,l}^{\textrm{self}}]^2 .
\end{align}
% The last one is a constant that stems from the self-interaction, which
% do not depends on the
Where
\begin{align}
  \v F_{\alpha,l} (\v r)
  &=
  \sum_{
    \begin{subarray}{c}
      \vert m_{\alpha}\vert < K_\alpha/2\\
      \v m\neq 0
    \end{subarray}}
  -4 \pi i\, l K_\alpha \v a^\ast_\alpha\,f(\v m)\,
  Z_{\alpha,l}(\v m)\,
  e^{2\pi i\v m\cdot \v r}.
  % \\
  % C_{\alpha,l}^{\textrm{self}}
  % &= 
  % \sum_{
  %   \begin{subarray}{c}
  %     \vert m_{\alpha}\vert < K_\alpha/2\\
  %     \v m\neq 0
  %   \end{subarray}}
  % (-2)f(\v m)\,
  % Z_{\alpha,l}(\v m).
\end{align}
% $F_{\alpha,l} (\v r)$ and $C_{\alpha,l}^{\textrm{self}}$ 
% do not depends on the coordinates of particles,
% so they can be tabulated in advance for the simulation.
Similar to $\v G_{\alpha,l}(\v r)$,
$\v F_{\alpha,l} (\v r)$ 
does not depends on the coordinates of particles,
so it can be tabulated  for the simulation.
The nearest neighbor approximation to the correlation error is
\begin{align}\nonumber
  \mathcal E^{\textrm{ana}}_{\textrm{correlation}}(\v r)
  =\,&
  q^2
  \big[\,
  \big\{
  \sum_{\alpha} \sum_{l\neq 0}
  (\v G_{\alpha,l}+\v F_{\alpha,l})\cdot
  [\hat T_{\textrm{O}} (\hat{\v G}_{\alpha,l}+\hat{\v F}_{\alpha,l})]^{\vee}
  \big\}
  \ast \rho_{\textrm{O}}
  \,\big] (\v r) \\\nonumber
  \,&+
  4q^2\,
  \big[
  \big(
  \sum_{\alpha} \sum_{l\neq 0}  
  \v G_{\alpha,l}
  \big)
  \cdot
  \big(
  \sum_{\alpha} \sum_{l\neq 0}  
  [\,\hat T_{\textrm{O}}\hat{\v G}_{\alpha,l}]^{\vee}
  \big)
  \ast \rho_{\textrm{O}}\,
  \big] (\v r) \\\nonumber
  \,&+
  q^2
  \big[\,
  \big\{
  \sum_{\alpha} \sum_{l\neq 0}
  (\v G_{\alpha,l}+\v F_{\alpha,l})\cdot
  [\hat T_{\textrm{H}} (\hat{\v G}_{\alpha,l}+\hat{\v F}_{\alpha,l})]^{\vee}
  \big\}
  \ast \rho_{\textrm{H}}
  \,\big] (\v r) \\
  \,&+
  4q^2\,
  \big[
  \big(
  \sum_{\alpha} \sum_{l\neq 0}  
  \v G_{\alpha,l}
  \big)
  \cdot
  \big(
  \sum_{\alpha} \sum_{l\neq 0}  
  [\,\hat T_{\textrm{H}}\hat{\v G}_{\alpha,l}]^{\vee}
  \big)
  \ast \rho_{\textrm{H}}\,
  \big] (\v r).
\end{align}
